---
title: "Lake Erie Transect Set Up"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lake Erie Transect Set Up}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

```{r setup}
library(erieacoustics)
```

## Introduction
The package `erieacoustics` provides tools to standardize the analysis of Lake Erie acoustic survey data across the three surveys (WB, CB, EB). Historically the surveys have followed similar analytical approaches as described in the Standard Operating Procedures this effort aims to provide a common suite of R tools to improve standardization as well as provide better documentation of the survey.


## Project Set Up
An RStudio project template has been created establish a common directory structure and necessary `README` files for the surveys. The common project directory structure is an important guiding principle as many of the functions expect data to be stored in a specific directory tree and common naming conventions. 

### Directory Structure
Folder | Description
---------- | --------------------------------------------------
1_Annual_Protocol | contains Rmarkdown scripts and products for annual survey protocol
2_EVTemplate | Echoview processing template should be stored here to be available to `set_up_transect` 
3_Ping_Data | contains the acoustic data for each transect
3_Ping_Data/TransectName | each transect should be uniquely named and all data stored within a single directory
4_Trawl_Data | contains the paired trawl data
5_Enviro_Data | contains environmental measurements
6_Misc | contains miscellaneous files associated with survey
7_Annual_Summary | contains Rmarkdown scripts and products for annual summary 

## Processing Steps

1. Check that a template exists
```{r}
evtemplate <- file.path(getwd(), '2_EVTemplate/ErieHacTemplate_2021.EV')
dir.exists("3_Ping_Data") 
file.exists(evtemplate)
```

2. Use COM scripting to open new file using the template, import all acoustic files (i.e. `*.raw`, `*.dt4`) and perform tasks within the Echoview Template.
```{r}
set_up_transect(evtemplate, projecthome = getwd(),
                sonartype = "BIOSONIC", tranname = "G_315_07062021")
```

3. View files with in EV and make necessary adjustments
<!-- insert some EV screen captures here -->

4. Define line relative regions
Important to follow the established naming convention. Provide a description and example here. Include a screen capture.

5. Export transect data
```{r}
export_transect_evdata(getwd(),"G_315_07062021",1000)
```

6. Run analysis scripts.


